---
title: "classIUPUI-vignette"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{classIUPUI-vignette}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(classIUPUI)
```

# Importing GEO data

The first step is to import the GEO data by entering the GEO ID you want. 

```{r}
# collects the gse object
gse <- import_GEO("GSE33126")

```


# Collect GSE Data

The next step is to extract the expression, annotation, and phenotype data from the GSE object.

```{r}

out_list <- GEO_data(gse)

phenotype_data <- out_list[['Phenotype_data']]

expression_data <- out_list[['Expression_data']]

annotation_data <- out_list[['Annotation_data']]

```

# Expression Data Summary and Normalization Via Box Plots

The next step is to check and normalize the expression data. The normalized data will be visualized in a box plot.

```{r}

expression_data <- check_Values(expression_data)

```
# Simplify Phenotype Data

The phenotype data contains multiple columns which we don't need so you can select the most important and rename them to human readable names.
```{r}

sample_info <- sample_Simplify(phenotype_data)

```

# Create Annotated Heatmap of Expression Data

The heatmap of the expression data can be annotated with sample_info from the simplified phenotype data. If you do not provide the sample info the heat map will be unannotated.

```{r}

# annotated heatmap
GEO_heatmap(expression_data,sample_info)

# unannotated heatmap
GEO_heatmap(expression_data)

```

# PCA Scatter Plot

Creates a scatter plot of the PCA vectors.

```{r}

PCA(expression_data,sample_info)

```

# Differential Expression with Limma

Uses the Limma library to preform a differential expression analysis on the gse object from import_GEO.

```{r}
# additional options include the number of threshold genes and the names for column 1 and 2
fit2 <- differential_Expression_limma(gse,sample_info,2,"col1","col2")

```

# Limma Data Annotation

We can add our annotation data to the limma fit2 data.

```{r}
# any top number of genes can be chosen
top_genes_frame <- annotate_fit(annotation_data,fit2,50)

```

# Generate Volcano Plot

With our fit2 data annotated we can now use it to generate a volcano plot.

```{r}
# any p-value cut off can be chosen for option 3
volcano_plot(top_genes_frame,fit2,0.05)

```

# Generate MA Plot

With our fit2 data annotated we can also  use it to generate a MA plot.

```{r}

MA_plot(fit2,top_genes_frame)

```

# Reactome Enrichment Anyalsis 

Using ReactomePA we can find the genes which are disproportionately enriched.

```{r}
# any p-value cut off can be chosen for option 3
x_enrichedPathway <- reactome_analysis_enrich(top_genes_frame,0.05)
print(x_enrichedPathway)

```
# GSE Pathway Analysis 

Using ReactomePA we can find the genes which are disproportionately enriched in the gene set pathway

```{r}

# any p-value cut off can be chosen for option 3
y_gsePathway <- reactome_analysis_gse(top_genes_frame,0.05)
print(y_gsePathway)
```

# Disease Set Enrichment Anaylsis 

We can similarly check for over-representation in disease ontology using DOSE.


```{r}
# any p-value cut off can be chosen for option 3
z_disease_enrichment <- disease_enrichment_analysis(top_genes_frame,0.05)
print(z_disease_enrichment)
```

